Reanalysis Step 9: Alternative Propensity Score Specifications

### 2025-09-05 reanalysis-9-alternative_specifications.qmd
table1 refactor

---
title: "Reanalysis Step 7: Alternative Propensity Score Specifications"
subtitle: "Comprehensive comparison of PS methods with multiple imputation"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
---

@@

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
setup_analysis(seed = 2025)

library(MatchIt)
library(WeightIt)
library(twang)
library(cobalt)
library(mice)
library(survey)
library(ggplot2)
library(gridExtra)
library(parallel)
library(doParallel)
library(forestplot)

@@

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))
ps_model_vars <- variable_lists$ps_model_vars

# Create PS formula
ps_formula <- as.formula(paste("ever_lapse_binary ~", 
                               paste(ps_model_vars, collapse = " + ")))

# ANALYSIS MODE
analysis_mode <- "final"  # Should match imputation mode
max_imputations_to_process <- ifelse(analysis_mode == "test", 5, n_imputations)

# Caching configuration
use_cache <- TRUE
cache_dir <- file.path(reanalysis_data_dir, "ps_alternative_methods")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
  log_message(paste0("Created cache directory: ", cache_dir))
}

# Parallel processing configuration
max_cores_for_ps <- 2  # Limit cores for memory management

log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Processing ", max_imputations_to_process, " of ", n_imputations, " imputations"))
log_message(paste0("Caching: ", ifelse(use_cache, "ENABLED", "DISABLED")))
```

## Define Propensity Score Methods

```{r define-methods}
# Define all PS methods to compare (excluding twang_gbm which we'll load from reanalysis-4)
ps_methods <- list(
  # MatchIt Methods - Nearest Neighbor with replacement
  list(name = "nearest_glm", package = "matchit", method = "nearest", 
       distance = "glm", estimand = "ATT", replace = TRUE),

  list(name = "nearest_gam", package = "matchit", method = "nearest", 
       distance = "gam", estimand = "ATT", replace = TRUE),

  list(name = "nearest_gbm", package = "matchit", method = "nearest", 
       distance = "gbm", estimand = "ATT", replace = TRUE,
       distance.options = list(n.trees = 1000, interaction.depth = 3)),

  list(name = "nearest_lasso", package = "matchit", method = "nearest", 
       distance = "lasso", estimand = "ATT", replace = TRUE),

  list(name = "nearest_rpart", package = "matchit", method = "nearest", 
       distance = "rpart", estimand = "ATT", replace = TRUE),

  list(name = "nearest_mahalanobis", package = "matchit", method = "nearest", 
       distance = "mahalanobis", estimand = "ATT", replace = TRUE),

  # MatchIt - Subclassification  
  list(name = "subclass_glm", package = "matchit", method = "subclass", 
       distance = "glm", estimand = "ATT", subclass = 5),

  # WeightIt Methods - IPTW
  list(name = "cbps", package = "weightit", method = "cbps", 
       estimand = "ATT", over = FALSE),

  list(name = "entropy", package = "weightit", method = "ebal", 
       estimand = "ATT"),

  list(name = "bart", package = "weightit", method = "bart", 
       estimand = "ATT")
)

# Note: twang_gbm results will be loaded from reanalysis-4

log_message(paste0("Defined ", length(ps_methods), " PS methods to compare"))
```

## Propensity Score Estimation Functions

```{r ps-functions}
# Function to process a single PS method for one imputation
process_ps_method <- function(imputation_id, data, method_params, ps_formula, cache_dir, use_cache = TRUE) {
  method_name <- method_params$name
  cache_file <- file.path(cache_dir, paste0(method_name, "_imp", imputation_id, ".rds"))

>>>>

---
title: "Reanalysis Step 9: Alternative Propensity Score Specifications"
subtitle: "Comprehensive comparison of PS methods with multiple imputation using strategy pattern"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
---

@@

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
source("alt_specs.R")  # Load strategy pattern implementation
setup_analysis(seed = 2025)

library(ggplot2)
library(knitr)

@@

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))

# ANALYSIS MODE: "test" or "final"
analysis_mode <- "final"  # Change to "test" for quick testing with 2 imputations

# Get configuration based on mode
config <- get_analysis_config(analysis_mode)

# Cache configuration
use_cache <- TRUE
cache_dir <- file.path(reanalysis_data_dir, "ps_alternative_methods")

log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Processing up to ", config$n_imputations, " of ", n_imputations, " available imputations"))
log_message(paste0("Caching: ", ifelse(use_cache, "ENABLED", "DISABLED")))
log_message(paste0("Cache directory: ", cache_dir))
```

## Define Formulas

```{r define-formulas}
# Get PS formula using reanalysis version (includes dr_severity)
ps_formula <- get_matching_formula_reanalysis()

# Get outcome formula
outcome_formulas <- get_analysis_formulas_reanalysis()
outcome_formula <- outcome_formulas$full

log_message("Propensity Score Formula:")
print(ps_formula)

log_message("Outcome Formula:")
print(outcome_formula)
```

## Available Methods

```{r show-methods}
# Get list of available methods
available_methods <- names(get_ps_methods(config))

cat("\nAvailable PS methods to run:\n")
for (method in available_methods) {
  cat(" -", method, "\n")
}

# Define which methods to actually run
# Comment out any methods you want to skip
methods_to_run <- c(
  "nearest_glm",
  "nearest_gam", 
  "nearest_gbm",
  "nearest_lasso",
  "nearest_rpart",
  "nearest_mahalanobis",
  "subclass_glm",
  "cbps",
  "entropy",
  "bart",
  "twang_gbm"  # Will load from reanalysis-4 if available
)

log_message(paste0("Will run ", length(methods_to_run), " methods"))
```

## Run Complete Analysis

```{r run-analysis}
# Run the complete analysis using the strategy pattern from alt_specs.R
# This will run all specified methods, fit outcome models, and pool results

results <- run_alternative_ps_analysis(
  imputed_datasets = imputed_datasets,
  ps_formula = ps_formula,
  outcome_formula = outcome_formula,
  mode = analysis_mode,
  cache_dir = cache_dir,
  use_cache = use_cache,
  methods_to_run = methods_to_run
)
